Verified Time Balancing of Security Protocols

### 2019-01-19 Src/talk.tex
more work on talk. Just need to describe Prg pictorially and then demonstrate syntax!

\title{Verified Time Balancing of Securit Protocols}
\author{Donovan Crichton}
\date{January 2019}

@@

\begin{frame}[fragile]
  \frametitle{The Problem}
  \begin{itemize}
    \item Can we formally prove that an implementation of
            a network security protocol is immune to 
                  timing side-channel attacks?
    \item In other words, can we show that a model of the protocol,
            along with its implementation is unable to leak
                  any timing information to an observer.
  \end{itemize}
\end{frame}

\begin{frame}[fragile]
  \frametitle{Rationale}
  \begin{itemize}
    \item Some departments in the ASD spend a large amount of time
            manually verifying cryptographic routines in vendor
            code.
    \item Requires large amounts of time and expertise, resulting
            in a slow verfication process.
    \item Formal methods may allow the 
            automation of these verification processes.
  \end{itemize}
\end{frame}

\begin{frame}[fragile]
  \frametitle{Formal Methods}
    \begin{itemize}
      \item A mathematical approach to the 
              development, specification and verification of 
                    software.
      \item Uses "theorem proving assistants", software that
              formally verifies against a specification.
      \item We can implement a proof of concept network security 
              protocol inside a theorem prover (namely Idris).
    \end{itemize}
\end{frame}

\begin{frame}[fragile]
  \frametitle{How do proof assistants work?}
    \begin{itemize}
      \item Rely on on a relationship between proofs in mathematics
              and computer programs known as the Curry-Howard 
                    correspondance.
      \item The proof system of intuitionistic 
              natural deduction can be directly interpreted 
                    as a model of computation known as 
                    lambda calculus.
    \end{itemize}
\end{frame}

\begin{frame}[fragile]
  \frametitle{A quick Idris example.}
  \begin{block}{A simple function in Idris}
  \begin{minted}{haskell}
f : Nat -> Nat
f x = x + 2
  \end{minted}
  \end{block}
  \begin{block}{The same function in familiar notation.}
$f : \mathbb{N} \mapsto \mathbb{N}$ \\
$f(x) = x + 2$
  \end{block}
\end{frame}

\begin{frame}[fragile]
 \frametitle{Logic in the Idris language.}
    \begin{table}[h!]
    \begin{tabular}{c|c|c|c}
    \textbf{Logic Term} & \textbf{Logic Symbol} & 
            \textbf{Idris Symbol} & \textbf{Idris Term} \\
    \hline
      Implication & p $\Rightarrow$ q & p \mintinline{haskell}{->} q
      & Arrow \\
      Conjunction & p $\land$ q & \mintinline{haskell}{(p, q)}
      & Pair (Product) \\ 
      Disjunction & p $\lor$ q & \mintinline{haskell}{Either p q}
      & Enum (Sum)\\
      Negation & $\lnot$ p & \mintinline{haskell}{p -> Void} &
      Void Type \\ 
      IFF/Eq & p $\equiv$ q, p $\iff$ q & \mintinline{haskell}{(p -> q, q -> p)}  
      & Pair Arrows \\
      Universal & $\forall$ x. P x &
      \mintinline{haskell}{p -> Type} & $\Pi$ Type \\
      Existential & $\exists$ x. P x
      & \mintinline{haskell}{(x ** P x)} & $\Sigma$ Type \\
      \hline
       & & \mintinline{haskell}{p = q} & Type Equality
    \end{tabular}
  \end{table}
\end{frame}

\begin{frame}[fragile]
  \frametitle{Outline of the problem.}
  \begin{itemize}
  \item It's much simpler to prove properties for a small
        language than a large one.
  \item We can then design a minimal programming language
        that is just powerful enough to express our example 
        protocol, while proving that this language has
        the properties we desire.
  \item We could then design a provably correct translation
        from this language into one more amenable to 
        production (C, Assembly, etc).
\end{frame}

\begin{frame}[fragile]
  \frametitle{Prg : A minimal core language}
  \begin{item} Every expression in this language is paramterised
          over its supposed running time. 
  \begin{item} The language can be elaborated into a language nicer
          to use.
  \begin{item} The language can also be compiled or translated
          into a language more amenable for real-life performance.
\end{frame}

\begin{frame}[fragile]
  \frametitle{A diagram of Prg.}
\end{frame}

\begin{frame}[fragile]
  \frametitle{Challenges and Outcomes}
  \begin{itemize}
    \item Still in the process of confirming whether Prf is 
            expressive enough for a cryptographic key exchange
    \item Are there other kinds of proofs we might be interested in
            on this language to ensure no additional leakage of
                  information?
    \item For future work, the implementation of the compilation
            or translation step in the thorem prover.
\end{frame}

%\begin{frame}[fragile]

%\end{frame}

>>>>

\title{Verified Time Balancing of Security Protocols}
\author{Donovan Crichton}
\date{January 2019}

@@

\begin{frame}[fragile]
  \frametitle{Motivation}
  \begin{itemize}
    \item ASD manually verifies vendor code with containing cryptographic
            processes.
    \item Formal Methods: A mathematically based approach to the
            specification and verification of software.
    \item Can we reduce some of the resources ASD spends on 
            manual verification by replacing with automatic verification?
    \item A case study.
  \end{itemize}
\end{frame}

\begin{frame}[fragile]
  \frametitle{Formally Verifying a Time Balanced Security Protocol}
  \begin{itemize}
    \item Attackers can gain information from message timing.
    \item Can we model this invariant? Naively all operations have the
            same running time.
    \item Can we ensure that assumptions on this model hold for the
            implementation?
  \end{itemize}
\end{frame}

\begin{frame}[fragile]
  \frametitle{The ZRTP Protocol}
  \begin{itemize}
    \item Initially started with ZRTP.
    \item ZRTP is complex and makes many decisions.
    \item Simplified version that contains just enough detail to 
            allow us to attempt to prove some interesting things!
  \end{itemize}
\end{frame}

\begin{frame}[fragile]
  \frametitle{The Simplified Protocol}
    \begin{itemize}
      \item Commit messages contain hashes of 256 bit random nonce.
      \item Diffie-Hellman key exchange contains modulo arithmetic.
      \item How can we formally guarantee the timing of operations?
    \end{itemize}
\end{frame}

\begin{frame}[fragile]
  \frametitle{Approach}
  \begin{itemize}
    \item We can use the notion of propositions as types.
    \item Dependently typed functional programming languages
            also act as theorem provers for a higher-order
                  constructive logic.
    \item We can model this protocol in such a language, Idris.
    \item We can express proofs about properties of the protocol.
  \end{itemize}
\end{frame}

\begin{frame}[fragile]
  \frametitle{Quick Background 1 - Currying}
  All functions treated as taking a single argument. \\ \\
  $f : \mathbb{N} \rightarrow \mathbb{N} \rightarrow \mathbb{N}$ \\
  $f = + $ \\ \\
  Applying an argument to a multi argument function returns the 
  rest of the function! \\ \\
  $f(2) : \mathbb{N} \rightarrow \mathbb{N}$ \\
  $f(2) = 2 + $ \\ \\
  Finally all arguments are applied. \\ \\
  $f(2, 3) : \mathbb{N}$ \\
  $f(2, 3) = 2 + 3 = 5 $
\end{frame}

\begin{frame}[fragile]
  \frametitle{Quick Background 2 - Propositions as Types}
     \begin{table}[h!]
    \begin{tabular}{c|c|c|c}
    \textbf{Logic Term} & \textbf{Logic Symbol} & \textbf{Idris Symbol}
      & \textbf{Idris Type} \\
    \hline
      Implication & p $\Rightarrow$ q & p \mintinline{haskell}{->} q
      & Arrow \\
      Conjunction & p $\land$ q & \mintinline{haskell}{(p, q)} 
      & Pair (Product) \\
      Disjunction & p $\lor$ q & \mintinline{haskell}{Either p q}
      & Enum (Sum)\\
      Negation & $\lnot$ p & \mintinline{haskell}{p -> Void} &
      Void Type \\
      IFF/Eq & p $\equiv$ q, p $\iff$ q & \mintinline{haskell}{(p -> q, q -> p)} 
      & Pair Arrows \\
      Universal & $\forall$ x. P x & 
      \mintinline{haskell}{p -> Type} & $\Pi$ Type \\
      Existential & $\exists$ x. P x 
      & \mintinline{haskell}{(x ** P x)} & $\Sigma$ Type \\
      \hline
       & & \mintinline{haskell}{p = q} & Type Equality
    \end{tabular}
  \end{table}
\end{frame}

\begin{frame}[fragile]
  \frametitle{Quick Background 3 - Idris Syntax and Functions as Types}
  Building a vector type in Idris: 
  \begin{minted}{haskell}
    data Vec : Nat -> Type -> Type where
      Nil  : Vec 0 a
      (::) : (x : a) -> Vec n a -> Vec (n + 1) a
  \end{minted}
  We can parameterise types over values to capture invariants in the model.
  \begin{minted}{haskell}
    append : Vec n a -> Vec m a -> Vec (n + m) a
    append Nil ys = ys
    append (x :: xs) ys = x :: append xs ys 
  \end{minted}
\end{frame}

\begin{frame}[fragile]
  \frametitle{Capturing time invariants in a small DSL}
  %insert picture for prg here!
\end{frame}
